Electron and ion acceleration by relativistic shocks: particle-in-cell simulations

### 2018-01-23 papers/Physica2017/Romansky_Physica2017.tex
paper

\begin{abstract}
	The problem of particle acceleration by collisionless astrophysical shocks is of fundamental importance for the cosmic ray origin problem and  for the high energy astrophysics as a whole. Fast magnetohydrodynamical shocks produced by relativistic jets are considered as very likely sites of cosmic ray acceleration.
	While cosmic ray acceleration by both non-relativistic and ultra relativistic shocks was studied in some details, the shocks of moderate Lorentz factors have attracted much less attention so far, while they are important for modeling gamma-ray burst afterglows and a number of other interesting objects. In this paper we present simulations of relativistic shock waves in electron-proton plasmas obtained with an implicit particle-in-cell code for the initial study the particle injection efficiencies.
\end{abstract}
\section{Introduction}
One of the possible sources of high-energy cosmic rays (CRs) are shock waves in astrophysical collisionless plasmas which are believed to accelerate particles via  the diffusive shock acceleration (DSA) mechanism, see, e. g. \cite{Bell1978}, \cite{Blandford1978}. The particle acceleration in the vicinity of shock front is a sophisticated and nonlinear problem, because the CRs strongly influence on the structure of the shock. Their pressure is comparable with the pressure of background plasma and significantly modifies the shock front (see, e.g., \cite{Bykov2014}). 
Also the instabilities in the anisotropic plasma can result to a strong magnetic field amplification. The spectra of synchrotron emission observed from supernova remnants (SNRs) indicate that the magnetic field in their shells is about 100 times stronger than the interstellar field (see, e.g., \cite{Berezhko2003},\cite{Uchiyama2007}). 
One of the most powerful methods to explore the processes in collisionless plasma is particle-in-cell (PIC) simulation.

@@

One can see in Figures \ref{protons} and \ref{electrons} the particle spectra for various inclination angles $\theta$. The spectra consist of three parts - the narrow peak of cold but fast moving initial flow, approximated by Maxwell distribution, shifted on the initial momentum, the wide peak of hot downstream, approximated by Maxwell-J{\"u}ttner distribution, and the non-thermal accelerated component. The small-scale oscilations of spectra are produced by numerical errors and should not be taken into account. Figures show that the spectrum of accelerated particles strongly depends on he inclination angle $\theta$. If $\theta$ is less than the critical value, defined by the equation $c\cdot \cos(\theta_{crit})=v_{shock}$, where all values are measured in the upstream rest frame, particles can escape from the front and cross it several times to gain more energy (see, e.g., \cite{Pelletier2017}), if $\theta$ is greater, shock becomes superluminal - particles need to move along the field lines faster than light to escape from the front, and acceleration is suppressed. It is difficult to evaluate the critical angle \textit{a priori}, because the compression relation of the shock which will be created in the simulation is unknown. However the estimate for the case of a strong wave (the compression relation equals $4$) provides $\theta_{crit}=45^{\circ}$ in the downstream frame, which is consistent with the results of our simulations.

@@

These results show some interesting effects for angles close to the critical value (black and yellow solid lines)- the electrons are still accelerating while the protons are not. It can be interpreted as electrons accelerate in the short-wavelength turbulence of the shock precursor. Due to the small gyroradii of the electrons in initial flow, the short-wavelength turbulence scatter them more efficiently than protons. However, further research is necessary to study this phenomenon in detail.

>>>>

\begin{abstract}
	The problem of particle acceleration by collisionless astrophysical shocks is of fundamental importance for the cosmic ray origin problem and  for the high energy astrophysics as a whole. Fast magnetohydrodynamical shocks produced by relativistic jets are considered as very likely sites of cosmic ray acceleration.
	While cosmic ray acceleration by both non-relativistic and ultra relativistic shocks was studied in some details, the shocks of moderate Lorentz factors have attracted much less attention so far, while they are important for modeling gamma-ray burst afterglows and a number of other interesting objects. In this paper we present simulations of relativistic shock waves in electron-proton plasmas obtained with an implicit particle-in-cell code for the initial study the particle injection efficiencies.
\end{abstract}
\section{Introduction}
One of the possible sources of high-energy cosmic rays (CRs) are shock waves in astrophysical collisionless plasmas which are believed to accelerate particles via  the diffusive shock acceleration (DSA) mechanism, see, e. g. \cite{Bell1978}, \cite{Blandford1978}. The particle acceleration in the vicinity of shock front is a sophisticated and nonlinear problem, because the CRs strongly influence on the structure of the shock. Their pressure is comparable with the pressure of background plasma and significantly modifies the shock front (see, e.g., \cite{Bykov2014}). 
Also the instabilities in the anisotropic plasma can result to a strong magnetic field amplification. The spectra of synchrotron emission observed from supernova remnants (SNRs) indicate that the magnetic field in their shells is about 100 times stronger than the interstellar field (see, e.g., \cite{Berezhko2003},\cite{Uchiyama2007}). 
One of the most powerful methods to explore the processes in collisionless plasma on microscopic scales is particle-in-cell (PIC) simulation.

@@

One can see in Figures \ref{protons} and \ref{electrons} the particle spectra for various inclination angles $\theta$. These spectra consist of three parts - the narrow peak of cold but fast moving initial flow, the wide peak of hot downstream, and the non-thermal accelerated component. Initial flow spectra are compared with Maxwell distribution, shifted on the initial momentum, and hot downstream spectra are compared with Maxwell-J{\"u}ttner distribution.  The small-scale oscilations of spectra are produced by numerical errors and should not be taken into account. Figures show that the spectrum of accelerated particles strongly depends on he inclination angle $\theta$. If $\theta$ is less than the critical value, defined by the equation $c\cdot \cos(\theta_{crit})=v_{shock}$, where all values are measured in the upstream rest frame, particles can escape from the front and cross it several times to gain more energy (see, e.g., \cite{Pelletier2017}), if $\theta$ is greater, shock becomes superluminal - particles need to move along the field lines faster than light to escape from the front, and acceleration is suppressed. It is difficult to evaluate the critical angle \textit{a priori}, because the compression relation of the shock which will be created in the simulation is unknown. However the estimate for the case of a strong wave (the compression relation equals $4$) provides $\theta_{crit}=45^{\circ}$ in the downstream frame, which is consistent with the results of our simulations.

@@

These results show some new interesting effects for angles close to the critical value (black and yellow solid lines)- the electrons are still accelerating while the protons are not. It can be interpreted as electrons accelerate in the short-wavelength turbulence of the shock precursor. Due to the small gyroradii of the electrons in initial flow, the short-wavelength turbulence scatter them more efficiently than protons. However, further research is necessary to study this phenomenon in detail.
